Formalizing Stack Safety as a Security Property

### 2023-01-21 submission5/core.tex
An attempt to classify stack security policies

>>>>

\rb{\textbf{A menagerie of protection mechanisms.}
%
We have made explicit the key point that there is \emph{not} a single stack
abstraction. This paves the way for a \emph{general} idea of stack protection
that is widely applicable with high-level concepts that remain unchanged (and I
think it's particularly important to at least introduce these more immutable
parts early on). The fact that this general ``framework'' requires low-level or
specific specializations is a predictable consequence of the protean nature of
the call stack.
%
\\
%
Let us return to a selection of representative policies in our menagerie and
consider how they fit into the framework. At the highest level, we can classify
all those techniques into two large groups, software-based and hardware-based.
Traditionally, most practical protections are implemented in software, so we
start by considering this first group.
%
\\
%
We can make two general observations about these software based techniques.
Firstly, (as we will see: discussion? technical details?) these techniques are
\emph{less modular} than their hardware counterparts. That is, the
countermeasures are tightly coupled with the program code they protect, and
consequently more effort is required to disentangle the program from the
protections, and to understand the impact of the latter on the former. This is
true of any systematic attempt to study the security of stack protection
policies, including their adaptation to our framework. (Owing to this added
complexity, software policies will not be the cleanest or most elegant exemplars
of use of the framework, its salient features and practical utility; this partly
motivates the focus on hardware policies.)
%
\\
%
(The above remark holds \emph{at least in a ``single-level model''} that
considers assembly language/machine code programs. In general, (low-level)
protections are introduced by a compiler or similar tool. One could consider a
multi-level extension of the framework pairing models of high-level programs and
the security transformations performed by a (stack-secure) compiler.)
%
\\
%
Secondly, the guarantees generally offered by software-based policies are
\emph{less absolute}, being defined more or less informally and usually
providing few definite guarantees of absence of \emph{general} classes of
errors. (The following discussion elaborates on this. It is conceivable that an
extended and more sophisticated framework could add \emph{probabilistic
reasoning} capabilities to capture and reason about the effects of less strict
protection mechanisms. This would be another orthogonal development to what is
presented here. In this paper, we are interested in assessing the complete
\emph{absence} of broad classes of vulnerabilities --- modulo any modeling gaps
or defects. In passing, this would extend the security properties to e.g. larger
system verification efforts.)
%
\\
%
One by one, consider the following specimens of software protection mechanisms
(quick explanations inlined?):
%
\\
%
\begin{itemize}
%
\item \emph{Stack canaries}. The purpose of canaries is to \emph{increase} the
  difficulty of some attacks on the stack, usually (always?) buffer overflows.
  They can never provide any absolute security guarantees and therefore do not
  satisfy any of our properties (some of the properties, e.g. confidentiality,
  are completely out of scope.) Although tedious, we could we could model this
  mechanism in our framework with the appropriate adaptations --- and
  unsurprisingly, testing would find property violations, which we already know
  are there: the results of this effort would be unsurprising and not as useful
  or instructive as other experiments.
%
\item \emph{Bounds checking}. This is fundamentally a high-level technique, even
  more obviously so than stack canaries (which could still detect violations
  caused by external code). Bounds checking only applies to code that has been
  instrumented by the ``secure compiler'' (i.e., to whole programs); it does not
  impose any constraints on adversarial code. The attacker model is therefore
  rather weak. (Certain variants, like SoftBound, are good targets for lifting
  our framework to the source level by e.g. using a C-language semantics and
  using Tagged C for security modeling; the lifted versions of our properties
  would likely hold under the expected assumptions. HardBound is probably an
  interesting transitional case because it uses capability-like hardware
  primitives for enforcement; it could be close to ongoing and future work to
  verify stack safety on capability machines, e.g. based on Cerise -- discuss
  together?)
%
\item \emph{Shadow stacks}. Like stack canaries, shadow stacks are
  ``probabilistic techniques'' in that they ``restrict the flexibility available
  in creating gadget chains'' (Shanbhogue et al.). This technique is solely
  concerned with control-flow integrity, and so closely related to our
  well-bracketed control flow. Nevertheless, and again, it is only an
  approximate technique, and our testing framework would be able to find the
  expected attacks against our security property.
%
\item \emph{Split stacks}. The security claims of split stacks are stronger than
  those of most software countermeasures, i.e., they are meant to guarantee
  their titular code-pointer integrity. They explicitly offer no protections on
  stack data (i.e., integrity and confidentiality), and so restrict themselves
  to well-bracketed control flow. (This is probably closer to our framework than
  most other software techniques, and it seems like it would be easier to model
  as well. However, this is again work based on high-level languages and would
  benefit from a higher level of abstraction.)
%
\item \ldots
%
\end{itemize}
%
Opposed to the above group of software techniques we have hardware-based
protection policies (more in background/technical preliminaries?). Although
programs must effectively leverage the protection mechanisms offered by the
hardware, the distinction between those protections and the program logic itself
is much cleaner: it is fairly simple to separate the use of the security
features in the hardware (and their effects in program code) from regular
program code. They also tend to aspire to offer complete protection against
entire classes of attacks. Both aspects make this family of techniques a better
fit for illustrating and showcasing our techniques.
%
\\
%
Why favor micro-policies? Their design is more flexible and arguably more
general than capability machines (of course, they pay a performance price for
this). This generality and ease of programming make them good candidates for an
exploration the design space of stack protection policies and proof of concept
implementations. On the practical side, there is a number of existing policies
available to examine (not just lazy policies!) --- and we find how they satisfy
our properties, how they do not, and errors in some of them, and see how to
correct them. Lazy enforcement is particularly novel, but not the only
possibility. This additional breadth is beneficial for the proof of concept
stage.
%
\\
%
Regarding capability machines, we know that most designs based on capability
machines do not satisfy our properties (or do so incurring significant costs),
and that fixes are not trivial and involve new types of capabilities (in a
sense, are they any more ``realistic?'' Only the latest Cerise paper would seem
to satisfy our properties, is directly influenced by previous versions of this
work, and is a reasonable target for ongoing/further study -- but this is not
easy, and even the author(s) estimate it would be a significant amount of work,
especially full proofs (references, communications, etc.).
%
\\
%
\ldots
%
}
